---
title: "R Tricks For Computational Stats"
author: Chenguang Pan
date: Jan 31, 2023
output:
  pdf_document:
    toc: false
    toc_depth: 4
    number_sections: false
    keep_tex: true
    highlight: tango
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = TRUE,
  prompt = TRUE,
  comment = '',
  collapse = TRUE,
  warning = FALSE,
  message = FALSE)
```

## 2.0 Week 2's in class notes

### 2.1 rnorm: generate data from normal distribution
`rnorm``to genenrate dataset from normal distribution.  
Using the `rnorm(n, mean =, sd=)` function, like
```{r}
sample_01 <- rnorm(n = 50, mean = 100, sd = 15)
var(sample_01)
# this sample's sd and mean is close to the population's
sd(sample_01)
mean(sample_01)
```   
Note it is not a vector. It is just `numeric` array.  

### 2.2 ECDF generate data from uniform and ECDF distribution
ECDF is empirical cumulative density function. For a given sample of observations from x1,x2,x3...xn and sort it from smallest to largest. The sample size is `n`. We give each datum a value of 1/n. The ECDF is to count the number of xi above a certain level.  
Note ECDF is a function, which returns the sum of i*(1/n) when x <xi. Here we use the `runif()` function to generate a dataset from uniform distribution. Note, here `runif()` is quiet like `rnorm()``
```{r,fig.show='hold',out.width="100%",out.height= "50%"}
set.seed(8289)
v1 <- runif(n=10, min = -1, max = 1)
v1
par(mfrow=c(2,2))
plot(ecdf(v1))
v2 <- runif(n = 100, min = -1, max = 1)
plot(ecdf(v2))
v3 <- runif(n = 1000, min = -1, max = 1)
plot(ecdf(v3))
```   

### 2.3 Generate data from exponential distribution   
TODO: to summarize every prob. distribution 
```{r,fig.show='hold',out.width="100%",out.height= "50%"}
set.seed(8289)
par(mfrow=c(2,2))
v1 <- rexp(n = 10, rate = .1)
plot(ecdf(v1))
v2 <- rexp(n = 100, rate = .1)
plot(ecdf(v2))
v3 <- rexp(n = 1000, rate = .1)
plot(ecdf(v3))
```   

### 2.4 Quantitle Function  
Quantile function is actually a inverse function of CDF. That is for a given probability, to find the greatest lower bound of value.  
Now, let's estimate the .25 quantile for a sample from uniform -1~1. Since the properties of ecdf is to give each datum a 1/n value and then to count the culmulative value of 1/n, and since this is a uniform distribution which means each two data point shares the same interval. Therefore the heigh change in step is 1/n   
```{r,fig.show='hold',out.width="100%",out.height= "50%"}
set.seed(9832)
v1 <- runif(n = 10, min= -1, max= 1)
# note ecdf is a function based on given observations
ecdf1<- ecdf(v1)
plot(ecdf1)
abline(a = 1/2, b = 1/2)
# note the following is not a quantile function!!
ecdf1(0.5)
```   
Next, let's estimate .5 quantile via ecdf. The quantile is greatest lower bound of set of x values for which the ecdf is greater or equal to .5.   
```{r, fig.show='hold',out.width="100%",out.height= "100%"}
# using cbind() funciton to combine tow columns of data
# the seq is to create a set of data with certain interval
ecdf_tab <- round(cbind(sort(v1), seq(.1,1,.1)),3)
colnames(ecdf_tab) <- c("x", "F(x)")
ecdf_tab
par(mfrow=c(1,2))
plot(ecdf_tab)
points(ecdf_tab[5,1], ecdf_tab[5,2], pch = 3, col = 2, cex = 2, lwd = 2)
# next, lets draw a quantile function using R's defualt
quantile(x = v1, prob = 0.5)
plot(ecdf_tab)
points(ecdf_tab[5,1], ecdf_tab[5,2], pch = 3, col = 2, cex = 2, lwd = 2)
points(quantile(x = v1, probs = .5), .5, pch = 4, col = 3, cex = 2, lwd = 2)
```     

By default, the `quantile()` function in R will return a different result. Here there are n = 10 observations.R uses type = 7 which specifies that the kth order statistic is assigned cumulative probability of (k-1)/(n-1). Furthermore, R help notes that "All sample quantiles are defined as weighted averages of consecutive order statistics."

```{r}
type7_tab <- round(cbind(sort(v1), seq(0, 1, 1/9)), 3)
colnames(type7_tab) <- c("x", "F(x)")
type7_tab
# Weighted avg of consecutive order statistics.
mean(type7_tab[5:6,1])
# the value is same to the quantile function
```

### 2.5 Multivariate Normal Distribution   
```{r,fig.show='hold',out.width="50%",out.height= "50%"}
# Generate multivariate normal
mvn_gen <- function(n, mu, sigma, factorization = "Cholesky") {
  # Generate a sample of size n from multivariate normal with
  # mean vector mu and covariance matrix sigma.
  # Argument factorization can be either "Cholesky" or "Spectral".
  d <- length(mu)
  Z <- matrix(rnorm(n*d), nrow = n, ncol = d)
  if(factorization == "Cholesky") { Q <- chol(sigma) } else {
    if(factorization == "Spectral") {
      ev <- eigen(sigma)
      lambda <- ev$values
      P <- ev$vectors
      Q <- P %*% diag(sqrt(lambda)) %*% t(P) } else {
        stop("Arg factorization must be 'Cholesky' or 'Spectral'.")
      } }
  mu <- matrix(mu, nrow = d, ncol = 1)
  J = matrix(1, nrow = n, ncol = 1)
  X <- Z %*% Q + J %*% t(mu)
  return(data.frame(X))
}

sig <- matrix(c(1, .5, .8, .5, 1, 0, .8, 0, 1), 3, 3, byrow = TRUE)

set.seed(2381)
mv1 <- mvn_gen(n = 1000, mu = c(0, 1, 4), sigma = sig, factorization = "Cholesky")
plot(mv1)
mv2 <- mvn_gen(n = 1000, mu = c(0, 1, 4), sigma = sig, factorization = "Spectral")
plot(mv2)
```

