---
title: "HUDM6026 Homework_02"
author: Chenguang Pan
date: Feb 03, 2023
output:
  pdf_document:
    toc: false
    toc_depth: 4
    number_sections: false
    keep_tex: true
    highlight: tango
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = TRUE,
  prompt = TRUE,
  comment = '',
  collapse = TRUE,
  warning = FALSE,
  message = FALSE)
```

## Question 01 SCR 3.3  
**MY SOLUTION:**  
The inverse transformation of the `Pareto(a,b)`'s cdf function is as followed. $$F^{-1}(u)=\frac{b}{(1-u)^\frac{1}{a}} $$  

```{r}
# define the quantile function of Pareto(a,b) distribution
quantile_Pareto <-function(prob, a, b){
  x <- b * (1-prob)^(-1/a)
  return(x)
}
# define the simulated sample size
n <- 100
u <- runif(n)
# based on the uniformly generated vector to get the random sample
X <- quantile_Pareto(u, 2, 2)
range(X)
```   
This inverse function runs well. Before comparing the simulated density and the original density, I derivate the CDF to get the pdf function of Pareto(a,b), that is:$$f(x)=\frac{ab^a}{x^{a+1}}$$
```{r,fig.show='hold',out.width="100%",out.height= "40%"}
# draw the density histogram of the simulated data
hist(X, prob = T, 
     breaks = 50, 
     main = expression(f(x)==ab^a/x^(a+1))) 
# prepare the Pareto(2,2) distribution
x <- seq(2,40,.38)
y <- 2*(2^2)/(x^(2+1))
# superimpose the lines on the simulated density
lines(x, y, col="red")
mtext("Figure 1. Comparing the simulated data with Pareto(a,b)",
      side = 3,
      line = -1,
      outer = T)
```   

## Question 02 SCR 3.9  
**MY SOLUTION:**  
This question has already given the clues to generate random variable for the rescaled Epanechnikov kernel
```{r,fig.show='hold',out.width="100%",out.height= "40%"}
# write a function based on text's information
gen_var <- function(n){ # n is the sample size
  U_1 <- runif(n, -1, 1)
  U_2 <- runif(n, -1, 1)
  U_3 <- runif(n, -1, 1)
  U_output <- c()
  for (i in c(1:n)) {
    if (abs(U_3[i]) > abs(U_2[i]) & 
        abs(U_3[i]) > abs(U_1[i]))
      {U_output[i] <- U_2[i]} 
    else 
      {U_output[i] <- U_3[i]}
  }
  return(U_output)
}

# generate 1000 data
U_output <- gen_var(1000)
hist(U_output, prob = T, 
     breaks = 100, 
     main = expression(f(x)==(3/4)*(1-x^2)))
x_vec <- seq(-1,1,0.001)
f_x <- 0.75*(1-x_vec^2)
lines(x_vec, f_x, col="red")
```

