---
title: "HUDM6026 Homework_07"
author: Chenguang Pan
date: Mar 17, 2023
output:
  pdf_document:
    toc: false
    toc_depth: 5
    number_sections: false
    keep_tex: true
    highlight: tang
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = TRUE,
  prompt = TRUE,
  comment = '',
  collapse = TRUE,
  warning = FALSE,
  message = FALSE)
```   

## Q1:
*Do some research on the two-sample Kolmogorov-Smirnov (K-S) test for equality of distributions. Describe the null and alternative hypotheses and discuss how the test statistic is computed*  
  
  
**MY SOLUTION:**  

### 1.1 Two sample K-S test's concept

The two-sample Kolmogorov-Smirnov(K-S) test is a nonparametric test used to test whether two datasets are from the same population. K-S test works by comparing the cumulative distribution functions (CDF) of the two datasets.

The null hypothesis is that two datasets are drawn from the same population or that they have the same underlying probability distributions. The alternative hypothesis is that they are not from the same population or that they do not have the same underlying probability distributions. That is, $$H_0: F_1(x)=F_2(x) \quad\text{ v.s. }\quad H_1: F_1(x)\neq F_2(x)$$, where $F_1(x)$ and $F_2(x)$ are the distributions of two sets of random variables $X_1$ and $X_2$.  

### 1.2 How to compute the two-sample K-S test  
The K-S test compares the largest vertical distance between the two empirical cumulative distribution functions(ECDF). This is also called the Kolmogorov-Smirnov statistic (D). The test statistic is defined as: $$D=max|F_1(x)-F_2(x)|$$, where $F_1(x)$ and $F_2(x)$ are the ECDF of two datasets being compared.   
The maximum distance (i.e., the $D$) between the two CDFs is then compared to a critical value $$c(\alpha)\cdot\sqrt{\frac{m+n}{m\cdot n}}$$, where $c(\alpha)= \sqrt{-0.5ln(0.5\alpha)}$, depending on the sample size and significance level chosen.



If the test statistic is greater than the critical value, the null hypothesis is rejected, indicating that the two sets of data are not drawn from the same population.  

## Q2:
*Install package MatchIt and load it. Then call data(lalonde). Examine the help on lalonde and describe the meaning of the treat and re78 variables.*  
  
  
**MY SOLUTION:**  
```{r, eval=FALSE}
install.packages("MatchIt")
```  
```{r}
# extract the basic information of the given dataset
library(MatchIt)
data(lalonde)
head(lalonde)
dim(lalonde)
```  
The help file shows that `lalonde` has one treated group from NSW with the size of 185 and one comparison group from PSID with the size of 429. The variable `treat` is the treatment assignment(1= treated, 0=control). The `re78` is income in 1978, in US dollars.  

## Q3:
*Run the two-sample K-S test to test if participant income in 1978 is identically distributed across treatment group assignment or not...*  
  
  
**MY SOLUTION:**  

```{r}
# subset the treatment and control groups
x <- lalonde$re78[lalonde$treat ==1]
y <- lalonde$re78[lalonde$treat ==0]

# run the two sample k-s test
ks.test(x, y)
```   
The test statistic $D$ is .0986. Since the p-value is .162 and it is greater than the significant level .05, we fail to reject the null hypothesis. That is, there is no statistically significant difference between the treatment and comparison groups.  

## Q4:
*Create a function that takes two arguments, x and y, each a vector of values, and outputs the value of the two-sample K-S statistic D*  
  
  
**MY SOLUTION:**   


