---
title: "HUDM6026 Homework_05"
author: Chenguang Pan & Seng Lei
date: Feb 24, 2023
output:
  pdf_document:
    toc: false
    toc_depth: 5
    number_sections: false
    keep_tex: true
    highlight: tang
---

```{r setup, include=FALSE}
## Global options
knitr::opts_chunk$set(
  cache = TRUE,
  prompt = TRUE,
  comment = '',
  collapse = TRUE,
  warning = FALSE,
  message = FALSE)
```

## Q1:
*Determine the first derivative of f and encode it in a function called f_prime.*  

**MY SOLUTION:**  
Based on chain rule, the first derivative of $f(x)$ is $$f(x)' = \frac{-2x}{x^2+1}+\frac{1}{3}x^{-\frac{2}{3}}$$. Based on this equation, I write the code below. Certainly, we can use the R-built-in function to get the derivative quickly.
```{r}
f_prime <- function(x) {
  out_ <- (-2*x)/(x^2 + 1) + (1/3)*(x^(-2/3))
  return(out_)
}
# since this question requires the maximum, I use -1 times the function.
f_prime_neg <- function(x){
  out_ <- (-2*x)/(x^2 + 1) + (1/3)*(x^(-2/3))
  out_ <- out_*(-1)
  return(out_)
}
```  

## Q2:
*Create a plot of f and f on [0,4] in different colors and line types and add a legend.* 
**MY SOLUTION:**  
```{r}
# write the original function with the name of f_
f_ <- function(x){
  out_ <- (-1)*log(x^2 + 1) + x^(1/3)
  return(out_)}
# since this question requires the maximum, I use -1 times the function.
f_neg <- function(x){
  out_ <- (-1)*log(x^2 + 1) + x^(1/3)
  return(out_*(-1))}

# first plot the original function
x <- seq(0,4,0.01)
# plot the original function with blue line
plot(x, f_(x), col="blue", type = "l", ylim = c(-1.5,1))
# plot the first derivative with red line
lines(x, f_prime(x), col="red", type = "l")
# add the legend
legend(3,1, inset = 0.1, c("f_","f_prime"), lty = 1, 
       col = c("blue","red"), title="line Type")
# add a horizontal line to indicate the y=0
abline(h=0,lty=3)
abline(v=0.3683525,lty=3)
mtext("Figure 1. The Lines of the Given Function and it's First Derivative",
      side = 3,
      line = -2,
      outer = T)
```

## Q3:
*Finish the functions that I started in the R code notes for univariate optimization for the golden section search, the bisection method, and Newton’s method.*   

**MY SOLUTION:**  
### 3.1 The Golden Section Search  
```{r}
golden <- function(f, int, precision = 1e-6)
{
  rho <- (3-sqrt(5))/2 # ::: Golden ratio
  # ::: Work out first iteration here
  f_a <- f(int[1] + rho*(diff(int)))
  f_b <- f(int[2] - rho*(diff(int)))
  ### How many iterations will we need to reach the desired precision?
  N <- ceiling(log(precision/(diff(int)))/log(1-rho))
  for (i in 1:(N))                    # index the number of iterations
  {
    if (f_a < f_b)  
    {
      int[2] <- int[2] - rho*(diff(int))
      f_b <- f(int[2])
    } else{
      if (f_a >= f_b)
      {
        int[1] <- int[1] + rho*(diff(int))
        f_a <- f(int[1])
      } }
  }
  int
  print(paste0("Iteration for ",N," times;",
               " The location is at ", round(int[1],6)))
}
```  

### 3.2 The Bisection Method  
More information about this method can be found on *Page 116* of Chong and Zak (2013).
```{r}
bisection <- function(f_prime, int, precision = 1e-7)
{
  # ::: f_prime is the function for the first derivative
  # ::: of f, int is an interval such as c(0,1) which 
  # ::: denotes the domain
  
  N <- ceiling(log(precision/(diff(int)))/log(.5))
  # find the midpoint of the initial uncertainty range
  midpoint <- (int[1]+int[2]) /2
  # evaluate the f_prime on the midpoint
  f_prime_a <- f_prime(midpoint)
  i <- 1
  for (i in 1:N)
  {
    i <- i + 1
    if(f_prime_a < 0)
    {
      # if the f_prime on the midpoint is less than 0,
      # the minimizer must be on the right side of midpoint.
      int[1] <- midpoint
      # update the uncertainty range
      midpoint <- (int[1]+int[2]) /2
    } else
      if(f_prime_a > 0)
      {
        # if the f_prime on the midpoint is less than 0,
        # the minimizer must be on the left side of midpoint.
        int[2] <- midpoint
        midpoint <- (int[1]+int[2]) /2
      } else
        if(f_prime_a == 0)
        {
          break
        }
    # ::: FILL IN CODE HERE (UPDATE)
    f_prime_a <- f_prime(midpoint)
  }
  int
  print(paste0("Iteration for ",N," times;",
               " The location is at ", round(int[1],6)))
}
```

### 3.3 The Newton's Method   
More information about this method can be found on *Page 116* of Chong and Zak (2013).
```{r}
newton <- function(f_prime, f_dbl, precision = 1e-6, start)
{
  x_old <- start
  x_new <- x_old - f_prime(x_old)/f_dbl(x_old)
  
  i <- 1
  print(paste0("Iteration ", i, "; Estimate = ", x_new) )
  while (abs(x_new-x_old) > precision){
    x_old <- x_new
    x_new <- x_old - f_prime(x_old)/f_dbl(x_old)
    # ::: redefine variables and calculate new estimate
    # ::: keep track of iteration history
    print(paste0("Iteration ", i+1, "; Estimate = ", x_new) )
    i <- i + 1
  }
  x_new
}
```  

## Q4:
*Apply each of the three functions to this example to discover the minimum. Keep track of and report the number of iterations required for each method. Report the coordinates of the minimum discovered by each of the three functions as well as the number of iterations required*   

**MY SOLUTION:** 

### 4.1 Apply the Golden Section Search
```{r}
golden(f_neg, c(0,4))
```   

### 4.2 Apply the Bisection Method
```{r}
bisection(f_prime_neg, c(0,4))
```  

### 4.3 Apply the Newton's Method  
First, I need to get the second derivative of the original function. Based on the Quotient Rule and Chain Rule, one can easily have $$f(x)''= -6x^2-2-\frac{2}{9}x^{-\frac{5}{3}}$$
```{r}
# write the second derivative function
f_dbl <- function(x){
  out_ <- -6*x^2-2-(2/9)*x^(-5/3)
  return(out_)
}
# since this question requires the maximum, I use -1 times the function.
f_dbl_neg <- function(x){
  out_ <- -6*x^2-2-(2/9)*x^(-5/3)
  return(out_*(-1))
}
```  
Plug the first and the second derivatives into the Newton's Method function.
```{r}
newton(f_prime_neg, f_dbl_neg,start = 1)
```  
All three methods have the identical results, that is the maximum point is at $x=.368352$. As for the iteration times at the x interval $[0,4]$, the Golden Section Search iterated 32 times; the Bisection Method iterated 26 times. However, Newton's method iteration time largely depends on the start numbers. That is, the more guessing number close to the target, the less times needed.




